# Dependencies

How and when to use code from other packages.

* Motivation
* Scoping: how functions find variables
* `NAMESPACE`
* `DESCRIPTION`
* When should you take a dependency?

## Live coding

Going to switch formats to live coding. 

I will regularly check my notes into Github. 

bit.ly/build-tt

`notes/4-dependencies.Rmd`

And you can also see the complete script. 
I'd recommend not looking at it today.

## Motivation

```{r}
sd
```

```{r}
x <- c(1, 1, 5, 9, 9)
sd(x)
```

Let's overwrite `var()` so it always returns 100
```{r}
var <- function(x, na.rm = FALSE){
  100
}
```

Poll: Now, what will we get when we run?
```{r}
sd(x)
```
* `4`, the same thing as before
* `10`, that's `sqrt(100)`
* I'm not sure


OK, what if we create our own sd function:
```{r}
my_sd <- function (x) {
  sqrt(var(x))
}
```

Poll: Now, what will we get when we run?
```{r}
my_sd(x)
```
* `4`, the same thing as before
* `10`, that's `sqrt(100)`
* I'm not sure

Start with "why the difference", 
then we'll talk about what this means for your packages.

### Your Turn: Poll

What will `f()` return?
```{r}
x <- 1
y <- 1
z <- 1
f <- function() {
  y <- 2
  z <- 2
  g <- function() {
    z <- 3
    c(x, y, z)
  }
  g()
}
f()
```

* 1, 1, 1
* 1, 1, 3
* 1, 2, 2
* 1, 2, 3

### Your Turn: Breakout Rooms

Discuss:  What does `f()` return?  
What are the rules R uses to find a variable?

Timed: 4 minutes

## Scoping

Scoping: rules that describe how R finds variables.

When you type `x`, R looks in the current environment,
if it doesn't find it, 
it looks in the parent environment, 
if it doesn't find it, 
it looks in the parent environment, 
and so on,
until it reaches the empty environment,
which would return an error.

This same rule is always applied, 
regardless of whether a function is in a package or script. 
So what's the difference? The environment of the function.

Helpful functions from rlang: `get_env`, `env_print`,  `env_parents()`

```{r}
library(rlang)
get_env(my_sd) 
env_print(get_env(my_sd))
# there is a var() in this environment, 
# (the one we defined)
# so that is what is used
```

```{r}
# the sd() function is in the `namespace:stats` environment
get_env(sd)
env_print(get_env(sd)) # lot's of bindings!
env_has(get_env(sd), "var") # one is var()

env_has(get_env(sd), "sqrt") # it doesn't have `sqrt()`
env_parents(get_env(sd)) # but one of its parents does
env_has(env_parents(get_env(sd))[[1]], "sqrt")
env_has(env_parents(get_env(sd))[[2]], "sqrt")
```

Let's look at the parents of some functions in packages:
```{r}
get_env(ggplot2::geom_point)
env_parents(get_env(ggplot2::geom_point))
```

```{r}
get_env(dplyr::mutate)
env_parents(get_env(dplyr::mutate))
```

Always the same. 
Functions used in a package function
will first be looked for in the package namespace
(i.e. all functions in the package),
then in the package imports,
then in the base package,
then in the global environment.

To use functions from other packages
in yours, 
you'll need to define the contents of the
import environemnt.

## Adding Dependencies to your Package

### How does this work in a package?

* `create_package("~/desktop/ns")`
* `use_mit_license()`
* `use_r("my_sd")`
    ```{r}
    #' My standard deviation function
    #'
    #' @param x A numeric vector
    #' @export
    my_sd <- function (x) {
      sqrt(var(x))
    }
    ```
* `check()`. 

```
> checking R code for possible problems ... NOTE
  my_sd: no visible global function definition for ‘var’
  Undefined global functions or variables:
    var
  Consider adding
    importFrom("stats", "var")
  to your NAMESPACE file.
```

**Your Turn** - Breakout Rooms

First, repeat the steps to illustrate the problem:
* `create_package("~/desktop/ns")`
* `use_mit_license()`
* `use_r("my_sd")`. Include the following function:
    ```{r}
    #' My standard deviation function
    #'
    #' @param x A numeric vector
    #' @export
    my_sd <- function (x) {
      sqrt(var(x))
    }
    ```
* `check()`. You should get a `NOTE` about `"Undefined global functions"`

Then, fix the problem. Include this in the Rxoygen comments
above the function:
```{r}
#' @importFrom stats var
```

`check()` to verify the `NOTE` is removed.

**Ready to move on?** Fill out the poll.

While you wait, open the `NAMESPACE` file, 
can you tell how it has changed?

### Questions?

### An Alternative Solution

Don't use `#' importFrom`, use `::` instead, e.g.
```{r}
#' My standard deviation function
#'
#' @param x A numeric vector
#' @export
my_sd <- function (x) {
  sqrt(stats::var(x))
}
```

Why doesn't `sqrt()` need a namespace or an import?
(Hint: what package is this found in? Why is it special?)

### Non-base packages

Things work a little differently (more obviously) when you're using a package in a non-base package:

```{r}
my_count <- function(x) {
  df <- tibble(x = x)
  count(df, x)
}
```

```{r}
check()
```

```
> checking R code for possible problems ... NOTE
  my_count: no visible global function definition for ‘tibble’
  my_count: no visible global function definition for ‘count’
```

Fix with `::`:
```{r}
my_count <- function(x) {
  df <- tibble::tibble(x = x)
  dplyr::count(df, x)
}
```

But you'll get a new `WARNING`  on `check()`:
```
> checking dependencies in R code ... WARNING
  '::' or ':::' imports not declared from:
    ‘dplyr’ ‘tibble’
```

This is a cryptic warning!  But, it indicates,
a package level problem, we haven't indicated
that a user needs `dplyr` and `tibble` installed,
for our code to work.

This change happens in the `DESCRIPTION`, easiest
with:
```{r}
usethis::use_package("dplyr")
usethis::use_package("tibble")
```

In `DESCRPITION`
```
...
Imports: 
    dplyr,
    tibble
```

Why didn't stats need this? Because it's a recommended package; so R knows that it's always going to be avaiable/installed.

* `DESCRIPTION` is all about package level. You need to maintain.
* `NAMESPACE` is at function level. Use documentation to maintain.

**Your Turn** Breakout Rooms

1. Add this function to your ns package
    ```{r}
    my_count <- function(x) {
      df <- tibble(x = x)
      count(df, x)
    }
    ```
2. Document the function

3. Get `check()` passing.

Ready to move on?  Check in with your group.
When everyone is ready, one person, answer the poll.

While you wait... re-write `my_count()` to use
the pipe `%>%`, then get `check()` passing.  
Hint: `?usethis::use_pipe()`

### Questions?

### Don't use `#' import`

E.g.
```{r}
#' @import stats
my_sd <- function (x) {
  sqrt(var(x))
}
```

Saves typing, but harder for you to keep track of where
functions are coming from.  More chance of conflicts.

### Recap of ways to use a function from another package

Always start with `use_package()`

* `::`, default, have to use for data.
* `@importFrom` common functions,  infix functions (like the pipe)
* `@import` only for packages specifically designed to be used in this way
* `foo <- pkg::foo` never do this unless you enjoy spending hours debugging.

## Exports

The search path that R uses to look for functions:
```{r}
search()
```

`#' @export` in Roxygen comment makes functions 
available when your package is loaded.  
You don't have to export all functions.

## When to take a dependency?

**Your turn**: Brainstorm the costs and benefits of taking a dependency.

Costs:
* install time
* download size
* system dependenices
* recursive dependencies
* behaviour of your package might change

Benefits:
* safer code
* faster code
* someone else wrote the code

Most of the costs are proportional to the number of people using your package. So when you start out, it's genuinely something that you shouldn't worry too much about. Keep an ear out for people having problems, and don't add gratuitiously, but otherwise don't worry about it. If you're using a single function from a package, might not be worth it. But beware false friends.
