# Tidyverse in packages

* Tidy evaluation
  * Data masking
  * Tidy select
* How to refer to functions (recap)
* Data in packages

Draws heavily on <https://dplyr.tidyverse.org/articles/programming.html>

I'm going to focus on ggplot2, and dplyr, 
but the same principles apply to basically any function in the tidyverse.

Evolved a lot over the last few years.  
Tidyverse has been so successful, 
because it emphasizes expression with less typing.

It does this a framework known as tidy evaluation.

Comes with some challenges,
you might encounter when you try to use
tidyverse functions inside your own functions.

## Data masking

Data masking is one kind of tidy evaluation - 
the framework that tidyverse packages use to make **doing** data analysis easier.

**Your Turn**: Breakout Rooms

The following calls use variables: 
`mtcars`, `cyl`, `mpg`, `starwars`, `homeworld`, and `species`.
But, you could group them into two types of variables. 

Discuss:  What are the two groups? How would you describe the groups?
```{r}
library(tidyverse) 

mtcars %>% 
  group_by(cyl) %>% 
  summarise(n = n(), mpg = mean(mpg))

starwars %>% 
  filter(homeworld == "Naboo", species == "Human")
```

Answer in:
slido.com #80875 
passcode: tidytools

---

* `mtcars`, `starwars` are "programming" variables 
    that exist in the environment, **env-variables**.  

*  `cyl`, `mpg`, `homeworld`, and `species` are "statistical" variables 
    that exist inside data, **data-variables**.

At its heart tidyeval blurs the line between 
variables in an environment (env-variables), and 
variables in a data frame (data-variables). 

### (+) Data-masking makes code more concise

This blurring gives much more concise code, 
e.g. compare to the base equivalent: 
```{r}
starwars[starwars$homeworld == "Naboo" & starwars$species == "Human", ]
```
Here we have to be explicit about `homeworld` and `species` being data-variables 
by prefacing with `starwars$`.

### (-) Data-masking makes some things harder

Two problems:

1. Can lead to ambiguity, subtle erros.  `R CMD check` can't find data-variables in the environment, so it complains. 

2. It's hard to use indirection; i.e. storing a data-variable in an env-variable

Two things to learn about:
* `.data` pronoun
* embracing with `{{`

### 1. Be explicit about data-vars and env-vars

```{r}
keep_small_cyl <- function(x){
  mtcars %>% 
    filter(cyl < x)
}
keep_small_cyl(5)
```

But there is a subtle problem, what if the data contains 
a column called `x`.

```{r}
mtcars$x <- 1
keep_small_cyl(5)
```

```
 mtcars %>% 
    filter(cyl < x)
```
`cyl` is a data-var, `x` is an env-var.  
rlang provides the `.data` and `.env` pronouns
to help disambiguate. 

Can use whenever an argument is indicated as `<data-masking>`.
+ `aes()` and `vars()` in ggplot2

```{r}
keep_small_cyl <- function(x){
  mtcars %>% 
    filter(.data$cyl < .env$x)
}
keep_small_cyl(5)
mtcars$x <- 1
keep_small_cyl(5)
```

**Your turn:** Code along (5 min)

Make these three functions refer explicitly to data-vars with `.data`:

1.
```{r}
cyl_plot <- function(){
  ggplot(mtcars) + 
    geom_bar(aes(cyl)) + 
    coord_flip()
}
cyl_plot()
```

2. 
```{r}
cyl_sum_wt <- function(){
  mtcars %>%
  group_by(cyl) %>%
  summarise(n = n(), avg = mean(wt))
}
cyl_sum_wt()
```

3.
```{r}
mpg_hist <- function(binwidth = 1){
  mtcars %>% 
    ggplot() +
    geom_histogram(aes(x = mpg), binwidth = binwidth)
}
mpg_hist()
```

**While you wait:** Can you extend `keep_small_cyl()` to work
with a user supplied string for the column name? E.g.

```{r, error = TRUE}
keep_small("hp", 65)
#>             mpg cyl  disp hp drat    wt  qsec vs am gear carb x
#> Merc 240D   24.4   4 146.7 62 3.69 3.190 20.00  1  0    4    2 1
#> Honda Civic 30.4   4  75.7 52 4.93 1.615 18.52  1  1    4    2 1

keep_small("mpg", 11)
#>                     mpg cyl disp  hp drat    wt  qsec vs am gear  carb x
#> Cadillac Fleetwood  10.4   8  472 205 2.93 5.250 17.98  0  0    3    4 1
#> Lincoln Continental 10.4   8  460 215 3.00 5.424 17.82  0  0    3    4 1
```

### 2. It's hard to use indirection

```{r}
cyl_plot <- function(){
  ggplot(mtcars) + 
    geom_bar(aes(cyl)) + 
    coord_flip()
}
```

Can we create `bar_plot()` that works like `cyl_plot()`,
except instead of always plotting the `cyl` column,
the user provides the variable to be plotted?

E.g. `bar_plot(cyl)` should give the same result as `cyl_plot()`

Try:
```{r}
bar_plot <- function(var){
  ggplot(mtcars) +
    geom_bar(aes(.data$var)) +
    coord_flip()
}
bar_plot(cyl)
```
``Error: Column `var` not found in `.data` ``

### 2. Fix by making user provide character string and use `.data`

Make user provide a string `bar_plot("cyl")`.
We can use `.data` but now provide `var` in `[[`:
```{r}
bar_plot <- function(var){
  ggplot(mtcars) +
    geom_bar(aes(.data[[var]])) +
    coord_flip()
}
bar_plot("cyl")
```

But what if you want the function to work like tidyverse functions: `bar_plot(cyl)`?

New idea! Embracing.

### 2. To work like tidyverse function, fix with embracing `{{`

Need to tell `aes()` that env-variable `var` contains a data-variable: 
do that with embracing, `{{`.

```{r}
bar_plot <- function(var){
  ggplot(mtcars) +
    geom_bar(aes({{ var }})) +
    coord_flip()
}
```

Think about `{{` as a special signal that this environment
variable contains a data variable.

Inspired by glue:
```{r}
name <- "Charlotte"
glue::glue('Hello, my name is { name }')
```


Other alternatives: make user provide `aes()` call, use `...`.

**Your turn**:  Breakout Rooms

Add embracing to make these three functions work:
1.
```{r}
# Summarise a column by cylinder
cyl_sum <- function(x){
  mtcars %>%
  group_by(.data$cyl) %>%
  summarise(n = n(), avg = mean(x))
}
cyl_sum(wt)
```
2.
```{r}
# A hexbin plot of two columns
hexbin <- function(df, x, y, bins = 10) {
  df %>% 
    ggplot(aes(x, y)) + 
    geom_hex(bins = bins)
}
hexbin(diamonds, carat, price)
```
3. 
```{r}
# three number summary of a column by a grouping variable
group_summary <- function(df, grp, x) {
  df %>% 
    group_by(grp) %>% 
    summarise(
      n = n(),
      min = min(x, na.rm = TRUE),
      max = max(x, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, mpg)
```
(8 mins)

**While you wait**: Allow `group_summary()` to take two grouping variables. 

---
**Solutions**

```{r}
group_summary <- function(df, grp, x) {
  df %>% 
    group_by({{ grp }}) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, mpg)
```

```{r}
group_summary <- function(df, grp, x) {
  df %>% 
    group_by(across({{ grp }})) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, c(cyl, carb), mpg)
```

### Data Masking - Summary

dplyr: `arrange()`, `filter()`, `group_by()`, `mutate()`, and `summarise()`  
ggplot2’s: `aes()`, `var()`

Arguments that use data masking 
will be indicated in docs with `<data-masking>`.

* You (as developer) know the data-variable is `x`: `arrange(df, .data$x)`
* Name of data-variable is a **character string** in an env-variable `x`: 
`arrange(df, .data[[x]])`
* Data-variable is passed as a function argument `x`: `arrange(df, {{ x }})`

Remember `...` can be passed through - no special treatment required.


## Tidy Selection 

The other half of tidy evaluation, used in functions like:
`select()`, `rename()`, `pivot_longer()` ...
When working with columns.

Implemented with tidyselect package: 
choose variables based on their position, name, or type.

Run into similar problem with indirection;
```{r}
my_select <- function(df, x) {
  df %>% 
    select(x)
}
my_select(mtcars, cyl)
```
<!-- Example fixed with embracing -->
```{r}
my_select <- function(df, x) {
  df %>% 
    select({{ x }})
}
my_select(mtcars, cyl)
my_select(mtcars, c(cyl, wt))
```

```r
vars <- c("cyl", "gear")
mtcars %>% 
  select(all_of(vars))
```

### Tidy select - Summary

Some functions have "select" semantics - 
you don't give them operations that work with values, 
but operations that work with column names:

* You (as developer) know the column is `x`:  `select(df, x)`
* Name of column is a **character string** in an env-variable `x`: 
`select(df, all_of(x))`
* Column is passed as a function argument `x`: `select(df, {{ x }})`

Remember `...` can be passed through - no special treatment required.

https://dplyr.tidyverse.org/articles/programming.html

Undergoing some changes for dplyr 1.0.0:

* `one_of()` replaced by `any_of()` and `all_of()` - 
what should happen if variable doesn't exist in data frame?

* Tidy selection inside data masking functions? New function `across()`, used with `mutate()` and `summarise()`,  replaces `mutate_at()`, `summarise_at()`, `mutate_if()` etc.

## Putting it all together

Get this package:
```{r}
usethis::create_from_github("cwickham/tidytest", fork = FALSE)
```

It has our fixed `cyl_sum()` function in it, 
but it isn't passing `check()`.

Your task: Get this package passing `check()`.

Hints:
* You'll need to make sure add the dplyr dependencies
* `.data` is from `rlang`, use
    ```{r}
    #' @importFrom rlang .data`
    ```
    and don't forget to add the dependency to your DESCRIPTION with
`use_package("rlang")`


## Data

To include data in a package:

* `use_data_raw("name_of_dataset")`
* creates an R file in `data-raw` with same name as your dataset
* `usethis::use_data()` to create the binary data file that R uses (wrapper around `save()`)

* To document: 
  * `use_r("data")`
  *
      ```R
      #' Title
      #'
      #' @description
      #' Data dictionary:
      #'
      #' * `var1`: description1
      "name_of_dataset"
      ```

A little more information at <https://roxygen2.r-lib.org/articles/rd.html#datasets>
