# Interface Part II: Output

In Interface I, talked about the parts of the interface that are 
Visible from the function signature (e.g what you see in Help under "Usage"): 
names and arguments/inputs.

Now, focus on the things that come out of a function: 
output, side effects and communication with the user.

```{r}
library(stringr)
```

* Type stability & Size stability
* Multiple outputs
* Side effects

---

## Warmup 

```{r, eval = FALSE}
y <- sum(x)
```

If this code runs (without error), 

* What can you tell me about `x`?
* What can you tell me about `y`?

Open ended poll

---

`x` is a numeric or logical vector (ignoring complex)
`y` is a numeric vector of length 1. (might be integer, might be double)

Two important properties: type and size.
---

## Type Stability

### Vector types

Today I'm going to focus on functions that work with vectors, because this covers most functions  that you might encounter in a data analysis. 

*Breakout Room* Brainstorm: What are the eight most common types of vectors? I.e. things you might find as columns in a dataframe/tibble.

---

logical, numeric, character, factor (ordered factor), date, 
time, duration, list.

(unimportant: raw, complex, integer/double distinction)

special: data.frame, NA

---

### Function types

The "type" of a function is:
the type of its inputs, and
the type of its output. 

Can you describe the type of these functions?

* `mean()`
* `diff()`
* `dplyr::first()`
* `paste()` (with `collapse`)

(You might also think about the size of the inputs and outputs)

---

* `mean()`
* `diff()`
* `dplyr::first()`
* `paste()` (with `collapse`)

### Types vs prototypes

`class()` is one way to think about type, 
but it isn't always that useful bec:

```{r}
class(1:10)
class(pi)
class("x")
```

But, you lose important attributes:
```{r} 
class(Sys.Date()) # what timezone?
class(factor("a")) # what levels?
```

Or the type of the contents:
```{r}
class(matrix(1:4, ncol = 2))
class(matrix(letters[1:4], ncol = 2))
```

The vctrs package introduces another way to think about type: a prototype.

This is a miniature version (zero observation version) of the vector that contains all of the attributes but none of the data.
```{r}
vec_ptype(1:10)
vec_ptype(factor("a"))
vec_ptype(Sys.Date())
vec_ptype(data.frame(x = 1))
vec_ptype_show(data.frame(x = 1))
vec_ptype_show(matrix(1:4, ncol = 2))
vec_ptype_show(matrix(letters[1:4], ncol = 2))
```

**Principle:** 

Functions are easy to reason about when their
output type is easily predicted from the input types.

**Type-stability:**  output type only depends on input type,
not input values.
 
```{r}
complicated_type <- function(x, y) {
  if (x < 0) {
    "negative"
  } else {
    if (is.na(y)) {
      NA
    } else {
      1
    }
  }
}
```

What happens if `x` is 
* numeric?
* logical?
* character?

A more realistic example:
```{r}
vec_ptype(median(1:3))
vec_ptype(median(1:4))
```

### `ifelse()`: Case study 

Predict the type of the output for these calls to `ifelse()`:
```{r}
ifelse(TRUE, 1, "no")
ifelse(FALSE, 1, "no")
ifelse(NA, 1, "no")

ifelse(TRUE, factor("a"), factor("b"))
ifelse(TRUE, Sys.Date(), Sys.Date() + 1)
```

Test your predictions with `vec_ptype_show()`

Is `ifelse()` type stable?

Take a look at the help for `ifelse()`.
What is the rule for the output type of `ifelse()`?

---
```{r}
vec_ptype_show(ifelse(TRUE, 1, "no"))
vec_ptype_show(ifelse(FALSE, 1, "no"))
vec_ptype_show(ifelse(NA, 1, "no"))

vec_ptype_show(ifelse(TRUE, factor("a"), factor("b")))
vec_ptype_show(ifelse(TRUE, Sys.Date(), Sys.Date() + 1))
```
---

```{r}
vec_ptype_show(NA)
vec_ptype_show(TRUE)
```

`ifelse()` is not type stable, since the output type depends on the values 
in `test`

```{r}
?ifelse
```

Basically, it's complicated.

### A better `ifelse()`

Breakout Room Brainstorm: 

* What conditions would you put on the types of the input?

* What rule would you use to determine the types of the output?

* How should the size of the output be determined?

---

The `dplyr::if_else(condition, true, false)` approach:

* `condition` must be logical, `true` and `false` must be the same type.
  `true` and `false` must be of length 1 or same length as `test`.

* type of output is type of `true` and `false`, attributes taken from `true`

* length of output is length of `condition`

It's type stable:
```{r}
vec_ptype_show(dplyr::if_else(TRUE, 1, 2))
vec_ptype_show(dplyr::if_else(FALSE, 1, 2))
vec_ptype_show(dplyr::if_else(NA, 1, 2))
```

But, the condition on same types can be annoying:
```{r}
dplyr::if_else(TRUE, "true", NA)
```

---

A better approach from: https://vctrs.r-lib.org/articles/stability.html#ifelse

* The first argument must be logical.

* `vec_ptype(if_else(test, yes, no))` equals `vec_ptype_common(yes, no)`. 

* `vec_size(if_else(test, yes, no))` equals `vec_size_common(test, yes, no)`. 

```{r}
if_else <- function(test, yes, no) {
  vec_assert(test, logical())
  c(yes, no) %<-% vec_cast_common(yes, no)
  c(test, yes, no) %<-% vec_recycle_common(test, yes, no)

  out <- vec_init(yes, vec_size(yes))
  vec_slice(out, test) <- vec_slice(yes, test)
  vec_slice(out, !test) <- vec_slice(no, !test)

  out
}
```

### Size-stability

Related idea:

A function is size-stable iff:

You can predict the output size knowing only the input sizes, or 
there is a single numeric input that specifies the output size.

Both size-stable and type-stable are examples of invariants: 
simple statements that describe the output of a function.

## Returning No Outputs

Usually a function that is called for its side effects:

* modifying files on disk
* displaying something to the user
* drawing a plot

### Don't combine computation and side effects in a single function

How can you extract the overall F-test p-value from `summary.lm()`?
```{r}
mod <- lm(mpg ~ wt, data = mtcars)
summary(mod)
```

You can't.

Hard to extract results of computation

Harder to analyse code because it may have non-local effects

### Return object invisibly

Allows use in pipelines.

E.g.
```{r}
library(readr)

mtcars %>% 
  write_csv("mtcars.csv") %>% 
  write_tsv("mtcars.tsv")
```

## Returning Multiple Outputs

Why does `str_locate()` return a matrix and not two vectors?

Why does `str_locate()` return a matrix and not a data frame? 

Why can `str_locate()` return a matrix while `str_locale_all()` has to return a list? 

```{r}
library(stringr)
x <- c(NA, "", "x", "y", "yy")
str_locate(x, "y")
str_locate_all(x, "y")
```
---

Same size: use a tibble, unless to work as input to other functions a matrix would be better.

Different sizes: use a list, name the elements.  If you are returning that same thing from many functions, make an S3 class.


## Other Output Considerations

Beyond vector functions:

* Output should feed easily into input

