# Object-oriented programming

* Base types - `typeof()`
* S3 - simplest system of OOP. Adding attributes to an object.
  The most important attribute `class`
* S4 - robust, but much more complicated, and unless you're working with Bioconductor. Don't use until there's a good book about it.
* R6 - very different to S3/S4, it's much more like OOP in python or javascript, or basically any other popular programming language of the last 20 years (apart from Julia). Powers reticulate. Useful in various places in Shiny. Avoided unless you really need it.

Five chapters: <https://adv-r.hadley.nz>.

## S3

The simplest thing that might possibly work. 
Built on top of regular R objects (atomic vectors, lists, functions, ...).

```{r}
df <- data.frame()
f <- factor()
mod <- lm(mpg ~ cyl, data = mtcars)

# You are already working with S3 objects
sloop::otype(df)
sloop::otype(f)
sloop::otype(mod)
sloop::otype(1:30)
sloop::otype(tibble::tibble())

# S3 object are built on base types
typeof(tibble::tibble())
typeof(df)
typeof(f)
typeof(mod)

# With additional attributes including class
attributes(df)
attributes(tibble::tibble())
attributes(f)
attributes(mod)
```

## Why are S3 objects useful?

Behaviour of a function can depend on the class:
```{r}
x <- 1:10
class(x)
summary(x)

x <- factor(1:10)
class(x)
typeof(x)
summary(x)
```

But it there is no protection, you can change the `class` attribute
and the behaviour will change:
```{r}
class(x) <- "Date"
print(x)
```

## Generics and methods

* `print()` is called a **generic function** = interface
* specific function is called **method** = implementation

To find out if a function is a generic: `sloop::ftype()`. 

The name of a method is just `{generic}.{class}`, and this is usually what you need to find help.

The details can get a bit more complicated so you can use `sloop::s3_dispatch()` to find it:
```{r}
library(sloop)
s3_dispatch(print(ordered(1:3)))
```

`=>`, `->` method exists and is found
* exists but isn't used 
Greyed out, doesn't exist

Use `s3_get_method()` to find the implementation.

## Your turn:

Use `sloop::s3_dispatch()` to find the print methods 
that are called for integers and dates.

```{r}
x <- c(1L, 2L)
y <- Sys.Date()
```

slido.com #80875 
passcode: tidytools

---

```{r}
sloop::s3_dispatch(print(x))
sloop::s3_get_method(print.default)
class(y)
sloop::s3_dispatch(print(y))
sloop::s3_get_method(print.Date)
```

Some examples where you might implement your own class in your package:

* You are returning complicated objects, like a list.
Let's you write a nice `print()` method,
to be easy on your users. Or separate computation from the printing side effect.

* You have a new data type you want to allow 
people to put in the columns of a tibble and have
a nice display.

* You have a function that returns a tibble, 
you want a pretty printing method, but you want
to still allow people to use the function in a dplyr pipeline.

## Create a new class by adding the class attribute

Let's practice by creating a new class based on functions,
called `secret`.  Won't display body of function, i.e. useful when teaching.

```{r}
f <- function(x) x * 100
f <- new_secret(f)
f
#> HIDDEN
f(1)
#> 100
```

What needs to happen inside `new_secret()`?




```{r}
new_secret <- function(f) {
  stopifnot(is.function(f))
  structure(f, class = "secret")
}
```

We want objects of class `secret` to have a special 
print method.  What function do we need to define?

* `print()`
* `print.secret()`
* `my_print()`
* `print_secret()`

slido.com #80875 
passcode: tidytools


Best practices at: https://adv-r.hadley.nz/s3.html#s3-classes
```{r}
sloop::s3_dispatch(print(y))
```

## Create a new method by creating a function with a special name

`{generic}.{class}`   
Arguments must match generic.

```{r}
print.secret <- function(x, ...) {
  cat("HIDDEN\n")
}

f <- function(x) x * 10
f2 <- new_secret(f)
f2
#> HIDDEN
f2(10)
#> [1] 100
```

### Case studies

**New vector data type:** Year-Month in tsibble

```{r}
# install.packages("tsibble")
library(tsibble)
x <- seq(as.Date("2016-01-01"), as.Date("2016-12-31"), by = "1 month")
year_mon <- yearmonth(x)
year_mon
tibble(year_mon = year_mon)
typeof(year_mon)
attributes(year_mon)
sloop::s3_dispatch(print(year_mon))
```

```{r}
year_mon
year_mon + 1
```

**Complicated object:** column specification from readr
```{r}
data <- read_csv(readr_example("mtcars.csv"))
View(spec(data))
spec(data)

s3_dispatch(print(spec(data)))
# most of the work happens in format.col_spec 
s3_get_method("print.col_spec")
s3_get_method("format.col_spec")
```

**Acts like a tibble**, looks like something else:
```{r}
# devtools::install_github("ropensci/skimr")
library(skimr)
dplyr::starwars %>% skim()
```

```{r}
dplyr::starwars %>% 
  skim() %>% 
  View()
dplyr::starwars %>% 
  skim() %>% 
  filter(n_missing > 0)
```

## Creating new generics: Motivation

We want a `bizarro()` function, that takes input and "bizarros" it:
```{r}
bizarro("abc")
#> [1] "cba"

bizarro(1)
#> [1] -1

bizarro(c(TRUE, FALSE))
#> [1] FALSE TRUE
```

```{r}
str_reverse <- function(x) {
  purrr::map_chr(stringr::str_split(x, ""), 
    ~ stringr::str_flatten(rev(.x))  
  )
}
```

Without a generic, `bizzaro()` function needs a ton of `if` clauses:
```{r}
bizarro <- function(x) {
  if (is.character(x)) {
    str_reverse(x)
  } else if (is.numeric(x)) {
    -x
  } else if (is.logical(x)) {
    !x
  } else {
    stop(
      "Don't know how to make bizzaro <", class(x)[[1]], ">", 
      call. = FALSE)
  }
}
```

Much better to create a `bizzaro` generic, then implement methods.

In interest of time skip rest of generics section.

* `create_package("~/desktop/bizzaro")`
* `use_mit_license()`
* `use_r("bizarro")`

```{r}
#' @export
bizarro <- function(x) {
  # This uses magic to pass arguments to the right method
  UseMethod("bizarro")
}

#' @export
bizarro.character <- function(x) {
  str_reverse(x)
}

str_reverse <- function(x) {
  purrr::map_chr(stringr::str_split(x, ""),
    ~ stringr::str_flatten(rev(.x))
  )
}
```

**Your turn**: Add:
* a numeric method that multiplies by -1
* a logical method which inverts TRUE/FALSE
* a data frame method that bizarros the column names, as well as each column.

Finish that? Write some documentation and some basic tests. Get check passing.

---

```{r}
bizarro.default <- function(x) {
  stop(
   "Don't know how to make bizzaro <",
   class(x)[[1]], ">",
   call. = FALSE
  )
}
```

```{r}
#' @export
bizarro.numeric <- function(x) {
  -x
}
```

```{r}
#' @export
bizarro.data.frame <- function(x) {
  x[] <- purrr::map(x, bizarro)
  setNames(x, bizarro(names(x)))
}
```

---

**Your turn**: Add a method for factors. You can use this test if you want.

```{r}
test_that("bizarro factors have levels reversed", {
  f1 <- factor(c("abc", "def", "abc"))
  f2 <- factor(c("cba", "fed", "cba"))

  expect_equal(bizarro(f1), f2)
  expect_equal(bizarro(f2), f1)
})
```

---

```{r}
#' @export
bizarro.factor <- function(x) {
  levels(x) <- bizarro(levels(x))
  x
}
```

## Conclusions

* Consider implementing your own S3 class if:
    * You return a list with a specific structure from many functions
    * You want a concise `print()` method for a complicated object
    
* Consider implementing your own generic if:
    * you want function behaviour to depend on the input type
    
