# Dependencies

How and when to use code from other packages.

* Motivation
* Scoping: how functions find variables
* `NAMESPACE`
* `DESCRIPTION`
* When should you take a dependency?


## Live coding

Going to switch formats to live coding. 

I will regularly check my notes into Github. 

bit.ly/build-tt

`notes/4-dependencies.Rmd`

## Motivation

```{r}
sd
```

```{r}
x <- c(1, 1, 5, 9, 9)
sd(x)
```

```{r}
var <- function(x, na.rm = FALSE){
  100
}
```

What will we get when we run?:
```{r}
sd(x)
```
Answer in:
slido.com #80875 
passcode: tidytools

```{r}
my_sd <- function(x){
  sqrt(var(x))
}
```

What will happen if we run:
```{r}
my_sd(x)
```
Answer in:
slido.com #80875 
passcode: tidytools

What will `f()` return?
```{r}
x <- 1
y <- 1
z <- 1
f <- function() {
  y <- 2
  z <- 2
  g <- function() {
    z <- 3
    c(x, y, z)
  }
  g()
}
f()
```
* 1, 1, 1
* 1, 1, 3
* 1, 2, 2
* 1, 2, 3

slido.com #80875 
passcode: tidytools

### Your Turn: Breakout Rooms

Discuss:  

* What does `f()` return?  
* What are the rules R uses to find a variable?

Timed: 4 minutes














