# Dependencies

How and when to use code from other packages.

* Motivation
* Scoping: how functions find variables
* `NAMESPACE`
* `DESCRIPTION`
* When should you take a dependency?


## Live coding

Going to switch formats to live coding. 

I will regularly check my notes into Github. 

bit.ly/build-tt

`notes/4-dependencies.Rmd`

## Motivation

```{r}
sd
```

```{r}
x <- c(1, 1, 5, 9, 9)
sd(x)
```

```{r}
var <- function(x, na.rm = FALSE){
  100
}
```

What will we get when we run?:
```{r}
sd(x)
```
Answer in:
slido.com #80875 
passcode: tidytools

```{r}
my_sd <- function(x){
  sqrt(var(x))
}
```

What will happen if we run:
```{r}
my_sd(x)
```
Answer in:
slido.com #80875 
passcode: tidytools

What will `f()` return?
```{r}
x <- 1
y <- 1
z <- 1
f <- function() {
  y <- 2
  z <- 2
  g <- function() {
    z <- 3
    c(x, y, z)
  }
  g()
}
f()
```
* 1, 1, 1
* 1, 1, 3
* 1, 2, 2
* 1, 2, 3

slido.com #80875 
passcode: tidytools

### Your Turn: Breakout Rooms

Discuss:  

* What does `f()` return?  
* What are the rules R uses to find a variable?

Timed: 4 minutes

What does `f` return?
```{r}
f()
```

## Scoping Rules

Rules that R uses to find variables.

R looks in the current environment first.
If R can't find it, R looks in the parent environment,
If R can't find it there, R looks in the parent environment,
and so, until it finds it, 
or it reach the empty environment.

The parent environment of a function is where it was 
defined (not called).

These same rules are followed for functions you've defined,
or for function inside packages.

useful functions from rlang:

```{r}
library(rlang)
get_env(my_sd)
env_print(get_env(my_sd))
# Q: for Charlotte why is parent rlang
```

```{r}
get_env(sd)
env_print(get_env(sd))
env_has(get_env(sd), "var")
env_parents(get_env(sd))
```
```{r}
get_env(ggplot2::geom_point)
env_parents(get_env(ggplot2::geom_point))
```
```{r}
get_env(dplyr::mutate)
env_parents(get_env(dplyr::mutate))
```

Always the same. 
Functions used in a package function
will first be looked for in the package namespace,
`namespace:package_name`
(i.e. all functions in the package),
then in the package imports, `imports:package_name`
then in the base package, 
then in the global environment.

To use functions from other packages
in yours, 
you'll need to define the contents of the
import environemnt.

## Adding Dependencies to your Package

**Your Turn** - Breakout Rooms

Get the ns package:
```{r}
usethis::create_from_github("cwickham/ns", fork = FALSE)
```

* `check()`. You should get a `NOTE` about `"Undefined global functions"`

* Fix the problem. Include this in the Roxygen comments
above the `my_sd()` function:
```{r}
#' @importFrom stats var
```

* `check()` to verify the `NOTE` is removed.

(~ 6 mins)

**Ready to move on?** Indicate in the Google Doc.

While you wait, open the `NAMESPACE` file, 
can you tell how it has changed?

### An Alternative Solution

Don't use `#' @importFrom`, use `::` instead, e.g.
```{r}
#' My standard deviation function
#'
#' @param x A numeric vector
#' @export
my_sd <- function (x) {
  sqrt(stats::var(x))
}
```

Why didn't we get a note for `sqrt()`? It's in the base
package.

### Non-base packages

Things work a little differently (more obviously) when you're using a package in a non-base package:

```{r}
my_count <- function(x) {
  df <- tibble(x = x)
  count(df, x)
}
```

```{r}
check()
```

```{r}
check()
```

```
> checking R code for possible problems ... NOTE
  my_count: no visible global function definition for ‘tibble’
  my_count: no visible global function definition for ‘count’
```

Fix with `::`:
```{r}
my_count <- function(x) {
  df <- tibble::tibble(x = x)
  dplyr::count(df, x)
}
```

But you'll get a new `WARNING`  on `check()`:
```
> checking dependencies in R code ... WARNING
  '::' or ':::' imports not declared from:
    ‘dplyr’ ‘tibble’
```

This is a cryptic warning!  But, it indicates,
a package level problem, we haven't indicated
that a user needs `dplyr` and `tibble` installed,
for our code to work.

This change happens in the `DESCRIPTION`, easiest
with:
```{r}
usethis::use_package("dplyr")
usethis::use_package("tibble")
```

In `DESCRPITION`
```
...
Imports: 
    dplyr,
    tibble
```

Why didn't stats need this? Because it's a recommended package; so R knows that it's always going to be avaiable/installed.

* `DESCRIPTION` is all about package level. You need to maintain.
* `NAMESPACE` is at function level. Use documentation to maintain.

**Your Turn** Breakout Rooms

1. Add this function to your ns package
    ```{r}
    my_count <- function(x) {
      df <- tibble(x = x)
      count(df, x)
    }
    ```
2. (Optional) Document the function

3. Get `check()` passing.
  * refer to functions explicitly `::`
  * `use_package()`

Ready to move on?  Check in with your group.
When everyone is ready, one person, answer the poll.

While you wait... re-write `my_count()` to use
the pipe `%>%`, then get `check()` passing.  
Hint: `?usethis::use_pipe()`

(5 mins)

### Don't use `#' import`

E.g.
```{r}
#' @import stats
my_sd <- function (x) {
  sqrt(var(x))
}
```

Saves typing, but harder for you to keep track of where
functions are coming from.  More chance of conflicts.

### Recap of ways to use a function from another package

Always start with `use_package()`

* `::`, default, have to use for data.
* `@importFrom` common functions,  infix functions (like the pipe)
* `@import` only for packages specifically designed to be used in this way
* `foo <- pkg::foo` never do this unless you enjoy spending hours debugging.

## Takeaways

Functions in packages are found using the same rules as
variables in R.

But, the environment of functions in packages
is the package namespace, then the package imports.

To use functions from other packages in your own, 
you need to:
* Be explicit about where the function is from, either:
    * `#' importFrom ` in Roxygen comment, or
    * use `::`
* Make sure your users get these package by including them in the `DESCRIPTION` file with:
    ```{r}
    usethis::use_package()
    ```
    
    
