# Tidyverse in packages

* Tidy evaluation
  * Data masking
  * Tidy select
* How to refer to functions (recap)

Draws heavily on <https://dplyr.tidyverse.org/articles/programming.html>

**Your Turn**: Breakout Rooms

The following calls use variables: 
`mtcars`, `cyl`, `mpg`, `starwars`, `homeworld`, and `species`.
But, you could group them into two types of variables. 

Discuss:  What are the two groups? How would you describe the groups?
```{r}
library(tidyverse) 
mtcars %>% 
  group_by(cyl) %>% 
  summarise(n = n(), mpg = mean(mpg))
starwars %>% 
  filter(homeworld == "Naboo", species == "Human")
```

Answer in:
slido.com #80875 
passcode: tidytools

(5-6 mins)

# Data Masking

One side to tidy evaluation.

## Two types of variables

* `mtcars` and `starwars`, datasets, *environmment variables* **env-vars**: exist in global environment.

* `mpg`, `homeworld`, `species`, columns/variables inside the data frames, *data variables* **data-vars** for short: exist inside a data frame / tibble.

Tidyverse blurs the distinction betweem the env-vars
and data-var, because it save typing, and makes life 
better:
```{r}
library(tidyverse)
starwars %>% 
  filter(homeworld == "Naboo", species == "Human")

starwars[starwars$homeworld == "Naboo" & starwars$species == "Human", ]
```


### (-) Data-masking makes some things harder

Two problems:

1. Can lead to ambiguity, subtle errors.  `R CMD check` can't find data-variables in the environment, so it complains. 

2. It's hard to use indirection; i.e. storing a data-variable in an env-variable

Two things to learn about:

* `.data` and `.env` pronouns
* embracing `{{` 

## Be explicit about where variables ocme from: .data or .env

```{r}
keep_small_cyl <- function(x){
  mtcars %>% 
    filter(cyl < x)
}
keep_small_cyl(5)
```
Subtle but important problem:
```{r}
mtcars$x <- 1
keep_small_cyl(5)
```
Ambiguity about where `x` should come from: env-variable, data-variable.

You have be explicit about data-vars and env-vars.
The package rlang provides `.data` and `.env`

```{r}
keep_small_cyl <- function(x){
  mtcars %>% 
    filter(.data$cyl < .env$x)
}
keep_small_cyl(5)
```

Can use whenever an argument is indicated as `<data-masking>`.
+ `aes()` and `vars()` in ggplot2

**Your turn:** Code along (5 min)

Make these three functions refer explicitly to data-vars with `.data`:

1.
```{r}
cyl_plot <- function(){
  ggplot(mtcars) + 
    geom_bar(aes(cyl)) + 
    coord_flip()
}
cyl_plot()
```

2. 
```{r}
cyl_sum_wt <- function(){
  mtcars %>%
  group_by(cyl) %>%
  summarise(n = n(), avg = mean(wt))
}
cyl_sum_wt()
```

3.
```{r}
mpg_hist <- function(binwidth = 1){
  mtcars %>% 
    ggplot() +
    geom_histogram(aes(x = mpg), binwidth = binwidth)
}
mpg_hist()
```

**While you wait:** Can you extend `keep_small_cyl()` to work
with a user supplied string for the column name? E.g.

```{r, error = TRUE}
keep_small("hp", 65)
#>             mpg cyl  disp hp drat    wt  qsec vs am gear carb x
#> Merc 240D   24.4   4 146.7 62 3.69 3.190 20.00  1  0    4    2 1
#> Honda Civic 30.4   4  75.7 52 4.93 1.615 18.52  1  1    4    2 1
keep_small("mpg", 11)
#>                     mpg cyl disp  hp drat    wt  qsec vs am gear  carb x
#> Cadillac Fleetwood  10.4   8  472 205 2.93 5.250 17.98  0  0    3    4 1
#> Lincoln Continental 10.4   8  460 215 3.00 5.424 17.82  0  0    3    4 1
```


