# Tidyverse in packages

* Tidy evaluation
  * Data masking
  * Tidy select
* How to refer to functions (recap)

Draws heavily on <https://dplyr.tidyverse.org/articles/programming.html>

**Your Turn**: Breakout Rooms

The following calls use variables: 
`mtcars`, `cyl`, `mpg`, `starwars`, `homeworld`, and `species`.
But, you could group them into two types of variables. 

Discuss:  What are the two groups? How would you describe the groups?
```{r}
library(tidyverse) 
mtcars %>% 
  group_by(cyl) %>% 
  summarise(n = n(), mpg = mean(mpg))
starwars %>% 
  filter(homeworld == "Naboo", species == "Human")
```

Answer in:
slido.com #80875 
passcode: tidytools

(5-6 mins)

# Data Masking

One side to tidy evaluation.

## Two types of variables

* `mtcars` and `starwars`, datasets, *environmment variables* **env-vars**: exist in global environment.

* `mpg`, `homeworld`, `species`, columns/variables inside the data frames, *data variables* **data-vars** for short: exist inside a data frame / tibble.

Tidyverse blurs the distinction betweem the env-vars
and data-var, because it save typing, and makes life 
better:
```{r}
library(tidyverse)
starwars %>% 
  filter(homeworld == "Naboo", species == "Human")

starwars[starwars$homeworld == "Naboo" & starwars$species == "Human", ]
```


### (-) Data-masking makes some things harder

Two problems:

1. Can lead to ambiguity, subtle errors.  `R CMD check` can't find data-variables in the environment, so it complains. 

2. It's hard to use indirection; i.e. storing a data-variable in an env-variable

Two things to learn about:

* `.data` and `.env` pronouns
* embracing `{{` 

## Be explicit about where variables ocme from: .data or .env

```{r}
keep_small_cyl <- function(x){
  mtcars %>% 
    filter(cyl < x)
}
keep_small_cyl(5)
```
Subtle but important problem:
```{r}
mtcars$x <- 1
keep_small_cyl(5)
```
Ambiguity about where `x` should come from: env-variable, data-variable.

You have be explicit about data-vars and env-vars.
The package rlang provides `.data` and `.env`

```{r}
keep_small_cyl <- function(x){
  mtcars %>% 
    filter(.data$cyl < .env$x)
}
keep_small_cyl(5)
```

Can use whenever an argument is indicated as `<data-masking>`.
+ `aes()` and `vars()` in ggplot2

**Your turn:** Code along (5 min)

Make these three functions refer explicitly to data-vars with `.data`:

1. `aes()` uses data masking
```{r}
cyl_plot <- function(){
  ggplot(mtcars) + 
    geom_bar(aes(.data$cyl)) +   
    coord_flip()
}
cyl_plot()
```

2. 
```{r}
cyl_sum_wt <- function(){
  mtcars %>%
  group_by(.data$cyl) %>%
  summarise(n = n(), avg = mean(.data$wt))
}
cyl_sum_wt()
```

3.
```{r}
mpg_hist <- function(binwidth = 1){
  mtcars %>% 
    ggplot() +
    geom_histogram(aes(x = .data$mpg), binwidth = binwidth)
}
mpg_hist()
```

**While you wait:** Can you extend `keep_small_cyl()` to work
with a user supplied string for the column name? E.g.

```{r, error = TRUE}
keep_small("hp", 65)
#>             mpg cyl  disp hp drat    wt  qsec vs am gear carb x
#> Merc 240D   24.4   4 146.7 62 3.69 3.190 20.00  1  0    4    2 1
#> Honda Civic 30.4   4  75.7 52 4.93 1.615 18.52  1  1    4    2 1
keep_small("mpg", 11)
#>                     mpg cyl disp  hp drat    wt  qsec vs am gear  carb x
#> Cadillac Fleetwood  10.4   8  472 205 2.93 5.250 17.98  0  0    3    4 1
#> Lincoln Continental 10.4   8  460 215 3.00 5.424 17.82  0  0    3    4 1
```

### 2. It's hard to use indirection
 
My goal: write function `bar_plot(cyl)` that gives same result as 
`cyl_plot()` but works for any column:

```{r}
bar_plot <- function(var){
  ggplot(mtcars) + 
    geom_bar(aes(.data$var)) + 
    coord_flip()
}
bar_plot(cyl)
```
```{r}
bar_plot <- function(var){
  ggplot(mtcars) + 
    geom_bar(aes(.data[[var]])) + 
    coord_flip()
}
bar_plot("cyl")
```
What we really want `bar_plot(cyl)`, or do things like `bar_plot(sqrt(cyl))`.

Embracing `{{`: Inside argument that are <data-masking>,
`{{ var }}` signal that `var` is an environment variable,
contains a data variable. 

`var` it's an argument, so inside the function its an
available variable.

But, it contain a data variable: `cyl`

```{r}
bar_plot <- function(var){
  ggplot(mtcars) + 
    geom_bar(aes({{ var }})) + 
    coord_flip()
}
bar_plot(gear)
```


**Your turn**:  Breakout Rooms

Add embracing to make these three functions work:
1.
```{r}
# Summarise a column by cylinder
cyl_sum <- function(x){
  mtcars %>%
  group_by(.data$cyl) %>%
  summarise(n = n(), avg = mean(x))
}
cyl_sum(wt)
```
2.
```{r}
# A hexbin plot of two columns
hexbin <- function(df, x, y, bins = 10) {
  df %>% 
    ggplot(aes(x, y)) + 
    geom_hex(bins = bins)
}
hexbin(diamonds, carat, price)
```
3. 
```{r}
# three number summary of a column by a grouping variable
group_summary <- function(df, grp, x) {
  df %>% 
    group_by(grp) %>% 
    summarise(
      n = n(),
      min = min(x, na.rm = TRUE),
      max = max(x, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, mpg)
```


**While you wait**: Allow `group_summary()` to take two grouping variables. 

**Your turn**:  Breakout Rooms

Add embracing to make these three functions work:
1.
```{r}
# Summarise a column by cylinder
cyl_sum <- function(x){
  mtcars %>%
  group_by(.data$cyl) %>%
  summarise(n = n(), avg = mean(x))
}
cyl_sum(wt)
```
2.
```{r}
# A hexbin plot of two columns
hexbin <- function(df, x, y, bins = 10) {
  df %>% 
    ggplot(aes(x, y)) + 
    geom_hex(bins = bins)
}
hexbin(diamonds, carat, price)
```
3. 
```{r}
# three number summary of a column by a grouping variable
group_summary <- function(df, grp, x) {
  df %>% 
    group_by(grp) %>% 
    summarise(
      n = n(),
      min = min(x, na.rm = TRUE),
      max = max(x, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, mpg)
```
(8 mins)

**While you wait**: Allow `group_summary()` to take two grouping variables. 

---
1.
```{r}
# Summarise a column by cylinder
cyl_sum <- function(x){
  mtcars %>%
  group_by(.data$cyl) %>%
  summarise(n = n(), avg = mean({{ x }}))
}
cyl_sum(gear)
```
2.
```{r}
# A hexbin plot of two columns
hexbin <- function(df, x, y, bins = 10) {
  df %>% 
    ggplot(aes({{ x }}, {{ y }})) + 
    geom_hex(bins = bins)
}
hexbin(diamonds, carat, price)
```
3. 
```{r}
# three number summary of a column by a grouping variable
group_summary <- function(df, grp, x) {
  df %>% 
    group_by({{ grp }}) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, mpg)
```
Allow `group_summary()` to take two grouping variables. 

Option 1: add another grouping argument

```{r}
# three number summary of a column by a grouping variable
group_summary <- function(df, grp1, grp2, x) {
  df %>% 
    group_by({{ grp1 }}, {{ grp2 }}) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, mpg, cyl, gear)
```

Option 2: Use the dots `...`

```{r}
# three number summary of a column by a grouping variable
group_summary <- function(df, x, ...) {
  df %>% 
    group_by(...) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, mpg, cyl, gear)
```
Option 3: Use `across()` to allow user to use tidyselect semantics
```{r}
# three number summary of a column by a grouping variable
group_summary <- function(df, grp, x) {
  df %>% 
    group_by(across({{ grp }})) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, c(cyl, gear), mpg)
group_summary(mtcars, starts_with("c"), mpg)
```


---

### Data Masking - Summary

dplyr: `arrange()`, `filter()`, `group_by()`, `mutate()`, and `summarise()`  
ggplot2’s: `aes()`, `var()`

Arguments that use data masking 
will be indicated in docs with `<data-masking>`.

* You (as developer) know the data-variable is `x`: `arrange(df, .data$x)`
* Name of data-variable is a **character string** in an env-variable `x`: 
`arrange(df, .data[[x]])`
* Data-variable is passed as a function argument `x`: `arrange(df, {{ x }})`

Remember `...` can be passed through - no special treatment required.

## Tidy Selection 

The other half of tidy evaluation, used in functions like:
`select()`, `rename()`, `pivot_longer()` ...
When working with columns.

Implemented with tidyselect package: 
choose variables based on their position, name, or type.

Run into similar problem with indirection;
```{r}
my_select <- function(df, x) {
  df %>% 
    select(x)
}
my_select(mtcars, cyl)
```
<!-- Example fixed with embracing -->
```{r}
my_select <- function(df, x) {
  df %>% 
    select({{ x }})
}
my_select(mtcars, cyl)
my_select(mtcars, c(cyl, wt))
```

* `all_of()` `any_of()` work with character strings
* `across()` uses tidyselect semantics and can be combined with `{{`

## Questions

Can we please learn a bit more about the benefit of `.data$cyl` vs. `mtcars$cyl`? 

`.data` will always refer to active data, e.g. `mtcars$cyl` wouldn't work here:
```{r}
cyl_plot <- function(){
  mtcars %>% 
    filter(gear == 4) %>% 
    ggplot() +
    geom_bar(aes(.data$cyl)) +   
    coord_flip()
}
cyl_plot()
```

## Putting it all together

Get this package:
```{r}
usethis::create_from_github("cwickham/tidytest", fork = FALSE)
```

It has our fixed `cyl_sum()` function in it, 
but it isn't passing `check()`.

Your task: Get this package passing `check()`.

Hints:
* You'll need to make sure add the dplyr dependencies
* `.data` is from `rlang`, use
    ```{r}
    #' @importFrom rlang .data`
    ```
    and don't forget to add the dependency to your DESCRIPTION with
`use_package("rlang")`



