# Notes

bit.ly/build-tt

`notes/6-interface.Rmd`

# Function interface

## Plan

* Intro to function interface
*
*

Goal: How do you make a package (functions) that is easy and pleasant to use?

* Interface: outside of the function. All the parts of a function that the user will see and interact with.
* Implementation: inside of the function. Specifics of how your algorithm is carried out.

Draws from: <http://design.tidyverse.org/>

## Your Turn: Poll

What are the properties of a function that define its interface? (Hint: There are at least 9!)

Time: 3 minutes

* Function name 
* Arguments (inputs)
  * Name
  * Types
  * Defaults 
* Output 
* Errors
* Side effects
  * Copy things to clipboard
  * Messages 
  * Save/write files
  
## Case study: base R string functions vs. stringr functions

Base R string functions are not wrong or bad! All functions can be improved, and it's useful to critique existing work. 

### Your Turn: Breakout rooms

Time: 10 minutes

Discussion:
* Critique the interface of the base R string functions.
  * Think particularly about names and arguments. 

Then:
* Answer the sli.do questions

```{r, eval = FALSE}
strsplit(x, split, ...)
grep(pattern, x, value = FALSE, ...)
grepl(pattern, x, ...)
sub(pattern, replacement, x, ...)
regexpr(pattern, text, ...)
gregexpr(pattern, text, ...)
substr(x, start, stop)
```

* Names
  * Kind of cryptic. Confusing. They are concise. 
  * No common theme. No common prefix. 
  * Difficult to distinguish. Hard to remember their differences. 

* Arguments
  * Names aren't consistent (sometimes `x`, sometimes `text`)
  * Order isn't consistent
  * Confusing
  
Possibility for error, also possibility for frustration and confusion. 

## Names

The name of a function a very important part of the interface.

* The first thing you interact with.
* The clue about what the function does. 
* Key to remembering what the function is supposed to do. 
* Looking up documentation. 

Guidelines: Function names should:

* Probably be verbs. 
* Be clear and concise. 
* Distinguish one function from related functions. 

### Your Turn: Breakout rooms

1. Match the following functions and arguments to their stringr equivalents.
2. What are some issues with the base R function names?
3. What are some issues with the stringr function names?

Evaluate on our three guidelines. 

Time: 8 minutes

```{r}
grepl(value = TRUE) # -> str_subset()
grep()              # -> str_which() 
grepl()             # -> str_detect()
sub()               # -> str_replace()
gsub()              # -> str_replace_all()
strsplit()          # -> str_split()
regexpr()           # -> str_locate()       
```

Base R function names:
* Not really verbs.
* Are concise, but they aren't very clear.
* Difficult to differentiate them. 

Stringr function names:
* Are verbs (for the most part).
* Clear and concise (for the part). Follow a common pattern, including helpful prefix.
* Somewhat easy to differentiate them. 

`str_detect()`, `str_which()`, and `str_subset()`. All the English words (detect, which, subset) have pretty similar meanings. 

```{r}
str_detect(c("apple", "pluot", "banana"), "a")
str_which(c("apple", "pluot", "banana"), "a")
str_subset(c("apple", "pluot", "banana"), "a")
```
What's the difference between `str_detect()`, `str_which()`, and `str_subset()`? Main difference is in their output type. 

```{r}
str_detect() # str_something_lgl() ?
str_which()  # str_something_int() ?
str_subset() # str_something_chr() ?
```

Verbs aren't always necessary. There are time that nouns can be a good idea. 

1. For example, when you're building up objects.

```{r}
# ggplot2 functions

library(ggplot2)

ggplot()
geom_histogram()
scale_x_discrete()
coord_flip()
```

2. When you're extracting a property.

```{r}
length()
str_length()

# get_length()
# str_get_length()
```

Guidelines: Function names should:

* Probably be verbs. 
* Be clear and concise. 
* Distinguish one function from related functions. 

### Prefixes and suffixes

* Use prefixes to group together related functions.
  Advantage: autocomplete
* Use suffixes for variations on a theme.
  * `str_replace()` and `str_replace_all()`
* Decide between `snake_case` and `camelCase`. Pick one and be consistent. 

### Guidelines for function names

Function names should:
* Probably be verbs.
* Be clear and concise. 
* Distinguish a function from related functions.
* (if relevant) use prefixes to group together related functions.
* (if relevant) use suffixes to identify functions that are variants of each other.

### Advice

* Try a thesaurus. 
* Choose dense words. 
* Edit down. There is other context besides just the function name. (`mutate()` doesn't need to be `mutate_data_frame()`).
* If you're stuck, try writing down exactly what your function does. 
* Consume good function names. 

## Arguments

* Types (groups)
* Properties of arguments:
  * Required / not required
  * Hidden / not hidden
  * Enumerated / not enumerated
  * Independent / dependent

### Your turn: Poll

Divide the arguments of the following functions into groups. 

```{r}
strsplit(x, split, fixed = TRUE)
grepl(pattern, x, fixed = TRUE, ignore.case = TRUE)
gregexpr(pattern, text, fixed = TRUE)
```

### Data, descriptors, and details

* Data: `x` and `text`

  * Primary vector of data. 
  * Required. 

* Descriptors: `pattern` and `split`

  * Essential feature of the operation to be carried out. 
  * Required.
  
* Details: `fixed` and `ignore.case`

  * Optional. 
  * Additional details. 
  
Order:

1. Data
2. Descriptors 
(3. `...`)
4. Details

Benefits:

* Works with the pipe ( %>% )
* Ordered from most important to least
* Forces details to be named

### Required

#### Your turn: Poll

Which of `sample()`'s arguments are required?

2 min

Simple rule:
* Required <-> no default
* Optional <-> has default

### Hidden

Avoid hidden arguments. 

Hidden arguments: when a function returns *surprisingly* different results on the same input. 

```{r}
tolower("I")
sort(letters)
```

```{r}
str_to_lower()
str_to_lower("I", locale = "tr")
```
```{r}
rnorm(1)
Sys.time()
```
Avoid hidden arguments. 

### Enums

```{r}
str_trim()
str_trunc()
rank()
```

Use if you have an arugment that takes only a small set of possible values.

*Key*: use `match.arg()`

```{r}
x <- function(animal = c("lion", "sloth")) {
  animal <- match.arg(animal)
  animal
}
```

### Independence 

Keep argument independent. 

Arguments are dependent if the possibilities for one depend on the values of another. 

```{r eval=FALSE}
# Can supply times and each, unless times is a vector
rep(1:3, times = 2, each = 3)
rep(1:3, times = 1:3, each = 3)
```
`times` and `each` are dependent.

Keep arguments independent. Having dependent arguments is more confusing for you, and it's also more confusing for the user. 

https://design.tidyverse.org/args-independence.html#how-do-i-remediate

### Recap

Names
* Use verbs.
* Be clear and concise.
* Think about how functions are told apart.
* Use prefixes and suffixes.

Arguments
* Data, descriptors, details. They should go in that order.
* Avoid hidden arguments
* Required arguments don't have defaults. Optional ones do.
* Keep arguments independent. 
* Use enums for arguments that can only take on one of several values.

# End-of-day survey

Please fill out on sli.do.
