## Interface II: Output

Yesterday: names, inputs to functions

Users see when they look at "Usage"

Output:
* return value +
* errors or messages 
* side effects (plot, printing on console, working with files on disk) +

## Warmup 

```{r, eval = FALSE}
y <- sum(x)
```

If this code runs (without error), 

* What can you tell me about `x`?
* What can you tell me about `y`?

slido.com #80875 
passcode: tidytools

* `x` is numeric (double and integers) or logical vector, no restriction on length
* `y` is numeric, always of length 1

```{r}
x <- c(1L, 2L)
y <- sum(x)
typeof(y)
```

Two important features of input and outputs:
* type 
* length

### Vector types

Brainstorm: What are the eight most common types of vectors? I.e. things you might find as columns in a dataframe/tibble.

logical, numeric, character, factor (ordered factor), date, 
time, duration, list.

(unimportant: raw, complex, integer/double distinction)

### Function types

The "type" of a function, is:
the type of its inputs, and 
the type of its output.

* `mean()` 
  * input: numeric vectors, logical vectors, Dates (?)
  * output: numeric, Date(?), length 1
  
* `paste(collapse = "")`
  * input: very few restrictions on type or length
  * output: character string, length 1
  
```{r}
x <- 3
paste(1:3, c("3", "3"), collapse = "")
```
  
### Types vs prototypes

One way to think about type is `class()`:
```{r}
class(1:10)
class(pi)
class("x")
```

You lose important attributes
```{r}
class(Sys.Date())
class(factor("a"))
```

```{r}
class(matrix(1:4, ncol = 2))
class(matrix(letters[1:4], ncol = 2))
```

Idea in the vctrs package: prototype (a more concrete defintion of type 
useful for vector types)

Miniature version (zero observation) of the vector:
```{r}
library(vctrs)
vec_ptype(1:10)
vec_ptype("hello")
vec_ptype(Sys.Date())
vec_ptype_show(Sys.Date()) # where is tz?(dates don't need timezone, duh?)

# better time zone example
class(Sys.time())
vec_ptype_show(Sys.time()) 
class(ISOdatetime(2020, 09, 17, 10, 59, 00, tz = "UTC"))
vec_ptype_show(ISOdatetime(2020, 09, 17, 10, 59, 00, tz = "UTC")) 

vec_ptype_show(matrix(1:4, ncol = 2))
vec_ptype_show(matrix(letters[1:4], ncol = 2))
```

**Principle:** 

Functions are easy to reason about when their
output type is easily predicted from the input types.

### Type stability

(principle the tidyverse is now adhering too)

**Type-stability**: Output type should depend only on the input types,
not input values. 

```{r}
complicated_type <- function(x, y) {
  if (x < 0) {
    "negative"
  } else {
    if (is.na(y)) {
      NA
    } else {
      1
    }
  }
}
```

What happens if `x` is?
* numeric, and negative output is character string
* numeric, and positive, it depends, but either `NA_logical`, numeric

```{r}
x <- 1:3
y <- 1:4
vec_ptype(x)
vec_ptype(y)
vec_ptype(median(x))
vec_ptype(median(y))
```

### `ifelse()`: Case study 

```{r}
x <- 1:4
x
ifelse(x < 2, yes = "small", no = "large")
```


Predict the type of the output for these calls to `ifelse()`:
```{r}
ifelse(TRUE, 1, "no")
ifelse(FALSE, 1, "no")
ifelse(NA, 1, "no")
ifelse(TRUE, factor("a"), factor("b"))
ifelse(TRUE, Sys.Date(), Sys.Date() + 1)
```

**Breakout Room**:

Test your predictions with `vec_ptype_show()`, `library(vctrs)`

Discuss: 
* Is `ifelse()` type stable?
* What is the rule for the output type of `ifelse()`? (Hint: `?ifelse()`)

Asnwer in Slido, once you are done with both questions: 
Is `ifelse()` type stable?

(5mins)

No, `ifelse()` is not type-stable:
```{r}
vec_ptype_show(ifelse(TRUE, 1, "no"))
vec_ptype_show(ifelse(FALSE, 1, "no"))
vec_ptype_show(ifelse(NA, 1, "no"))
vec_ptype_show(NA)
vec_ptype_show(TRUE)
```

Not a problem for type stability, but does make it
hard to predict output type:
```{r}
vec_ptype_show(ifelse(TRUE, factor("a"), factor("b")))
vec_ptype_show(ifelse(TRUE, Sys.Date(), Sys.Date() + 1))
```

* What is the rule for the output type of `ifelse()`? (Hint: `?ifelse()`) 

It is complicated.  

### A better `ifelse()`

**Breakout Room** (5mins)

Brainstorm: 

* What conditions would you put on the types of the input?

* What rule would you use to determine the types of the output?

* (Optional) How should the size of the output be determined?

