# Object Oriented Programming / S3

* Base types - `typeof()` 
* S3 - simplest thing that might work
* S4 - complicated, stricter, Bioconductor
* R6 - more similar to other OO languages

Five chapters Advanced R Book

## S3

Simplest thing that might work
You are alreaady working with s3 objects

```{r}
df <- data.frame()
f <- factor()
mod <- lm(mpg ~ cyl, data = mtcars)
```

You're alreayd working with S3 objects:
```{r}
library(sloop)
otype(df)
otype(f)
otype(mod)
```

All S3 objects are built on a base type, get this with `typeof()`
```{r}
typeof(df)
typeof(f)
typeof(mod)
```

With the addition of attributes
```{r}
attributes(df)
attributes(f)
attributes(mod)
```

`class` attribute, that makes S3 work.

## Why are S3 object useful?

Function behaviour depend on the class:
```{r}
summary(mod)
df <- data.frame(x = rnorm(10))
summary(df)
f <- factor(1:10)
print(f)
f
summary(f)
```

No protection, you change the `class` attribute and now functions
will act like the object is of the new class:
```{r}
f
class(f) <- "Date"
print(f)
summary(f)
```

## Generics and methods 

* `summary()`, `print()` are known as "generic functions" or "generics" for short = interface of the function.

* specific function that is called, is called a "method" = implementation

`sloop::ftype()`
```{r}
ftype(print)
ftype(plot)
ftype(mean)
# Charlotte think abotu functions that aren't
```

A method follows a naming convention: {generic}.{class}
`summary.lm` is the method of `summary()` for objects of class `lm`.

`sloop::s3_dispatch()`
```{r}
o <- ordered(1:3)
print(o) 
class(o) # guess it's looking for `print.ordered()`
s3_dispatch(print(o))
```

```{r}
print.factor
```
## Your turn:

Use `sloop::s3_dispatch()` to find the `print()` methods 
that are called for integers and dates.

```{r}
x <- c(1L, 2L)
y <- Sys.Date()
```

slido.com #80875 
passcode: tidytools

```{r}
countdown::countdown_fullscreen(3)
```

```{r}
s3_dispatch(print(x))
s3_get_method(print.default)
```

```{r}
s3_dispatch(print(y))
class(y)
s3_get_method(print.Date)
```
## Why would I make my own class?

Some examples:

* You are returning complicated objects from your functions.  You can write a `print()` method that looks nice. `str(mod)`, `mod`

* You have a new vector data type that you want to put in data frames/tibbles.  

* You have a function taht returns something like tibble, but you want a nicer/different/special display.

## Let's make a new class!

* base type: function
* class: secret

```{r}
f <- function(x) x * 100
f <- new_secret(f)
f
#> HIDDEN
f(1)
#> 100
```

Slido:
What needs to happen inside `new_secret()` to make `f` an
object of class `"secret"`?

```{r}
new_secret <- function(f){
  stopifnot(is.function(f))
  structure(f, class = "secret")
}

f <- function(x) x * 100
f <- new_secret(f)
f(1)
```

We want objects of class `secret` to have a special 
print method.  What function do we need to define?

```{r}
f
print.secret <- function(x, ...) {
  cat("HIDDEN\n")
}
f
s3_dispatch(print(f))
```

```{r}
f <- function(x) x * 100
f <- new_secret(f)
f
#> HIDDEN
f(1)
#> 100
```

## Case Studies

**New vector data type:** Year-Month in tsibble

```{r}
# install.packages("tsibble")
library(tsibble)
x <- seq(as.Date("2016-01-01"), as.Date("2016-12-31"), by = "1 month")
year_mon <- yearmonth(x)
year_mon
tibble(year_mon = year_mon)
typeof(year_mon)
attributes(year_mon)
sloop::s3_dispatch(print(year_mon))
```

```{r}
year_mon
year_mon + 1
```

**Complicated object:** column specification from readr
```{r}
library(tidyverse)
data <- read_csv(readr_example("mtcars.csv"))
View(spec(data))
spec(data)
s3_dispatch(print(spec(data)))
# most of the work happens in format.col_spec 
s3_get_method("print.col_spec")
s3_get_method("format.col_spec")
```
**Acts like a tibble**, looks like something else:
```{r}
# devtools::install_github("ropensci/skimr")
library(skimr)
dplyr::starwars %>% skim()
```

```{r}
dplyr::starwars %>% 
  skim() %>% 
  View()
dplyr::starwars %>% 
  skim() %>% 
  filter(n_missing > 0)
```

## What would I make my own generic?

We want a `bizarro()` function, that takes input and "bizarros" it:
```{r}
bizarro("abc")
#> [1] "cba"
bizarro(1)
#> [1] -1
bizarro(c(TRUE, FALSE))
#> [1] FALSE TRUE
```

```{r}
str_reverse <- function(x) {
  purrr::map_chr(stringr::str_split(x, ""), 
    ~ stringr::str_flatten(rev(.x))  
  )
}
```

Without a generic, `bizzaro()` function needs a ton of `if` clauses:
```{r}
bizarro <- function(x) {
  if (is.character(x)) {
    str_reverse(x)
  } else if (is.numeric(x)) {
    -x
  } else if (is.logical(x)) {
    !x
  } else {
    stop(
      "Don't know how to make bizzaro <", class(x)[[1]], ">", 
      call. = FALSE)
  }
}
```

## Conclusions

* Consider implementing your own S3 class if:
    * You return a list with a specific structure from many functions
    * You want a concise `print()` method for a complicated object
    
* Consider implementing your own generic if:
    * you want function behaviour to depend on the input type
