# Object Oriented Programming / S3

* Base types - `typeof()` 
* S3 - simplest thing that might work
* S4 - complicated, stricter, Bioconductor
* R6 - more similar to other OO languages

Five chapters Advanced R Book

## S3

Simplest thing that might work
You are alreaady working with s3 objects

```{r}
df <- data.frame()
f <- factor()
mod <- lm(mpg ~ cyl, data = mtcars)
```

You're alreayd working with S3 objects:
```{r}
library(sloop)
otype(df)
otype(f)
otype(mod)
```

All S3 objects are built on a base type, get this with `typeof()`
```{r}
typeof(df)
typeof(f)
typeof(mod)
```

With the addition of attributes
```{r}
attributes(df)
attributes(f)
attributes(mod)
```

`class` attribute, that makes S3 work.

## Why are S3 object useful?

Function behaviour depend on the class:
```{r}
summary(mod)
df <- data.frame(x = rnorm(10))
summary(df)
f <- factor(1:10)
print(f)
f
summary(f)
```

No protection, you change the `class` attribute and now functions
will act like the object is of the new class:
```{r}
f
class(f) <- "Date"
print(f)
summary(f)
```

## Generics and methods 

* `summary()`, `print()` are known as "generic functions" or "generics" for short = interface of the function.

* specific function that is called, is called a "method" = implementation

`sloop::ftype()`
```{r}
ftype(print)
ftype(plot)
ftype(mean)
# Charlotte think abotu functions that aren't
```

A method follows a naming convention: {generic}.{class}
`summary.lm` is the method of `summary()` for objects of class `lm`.

`sloop::s3_dispatch()`
```{r}
o <- ordered(1:3)
print(o) 
class(o) # guess it's looking for `print.ordered()`
s3_dispatch(print(o))
```

```{r}
print.factor
```
## Your turn:

Use `sloop::s3_dispatch()` to find the `print()` methods 
that are called for integers and dates.

```{r}
x <- c(1L, 2L)
y <- Sys.Date()
```

slido.com #80875 
passcode: tidytools






